---
title: "Experiment 1 Analysis"
output: 
  html_document:
    toc: true
---



**Sean Trott and Benjamin Bergen** 
2019

# Data reading, cleaning

```{r include=FALSE}
library(tidyverse)
library(forcats)
library(lme4)
library(effects)
library(psych)
```



```{r}
# setwd("/Users/seantrott/Dropbox/UCSD/Research/IndirectSpeechActs/2019_mentalizing_data/Experiment1/")
main_plus_sst = read_csv("data/exp1_aggregated.csv")

```

## Overall demographics

```{r}
table(distinct(main_plus_sst, subject, gender)$gender)
mean(distinct(main_plus_sst, subject, age)$age)
sd(distinct(main_plus_sst, subject, age)$age)
range(distinct(main_plus_sst)$age)

```

Differences in demographics across experiments?

```{r}

only_unique_subjects = main_plus_sst %>%
  distinct(subject, gender, assignment, age)

table(only_unique_subjects$assignment)

summary(aov(data = only_unique_subjects,
           age ~ assignment))


df_with_gender = only_unique_subjects %>%
  group_by(assignment, gender) %>%
  summarise(num_gender = n())

M = as.table(rbind(filter(df_with_gender, assignment == "4a")$num_gender,
             rbind(filter(df_with_gender, assignment == "4b")$num_gender)))

dimnames(M) <- list(gender = c("Female", "Male"),
                    assignment = c("4a", "4b"))

M
chisq.test(M)

```


## Break into two experiments

```{r}
coded_3a = main_plus_sst %>%
  filter(assignment == "4a")
length(unique(coded_3a$subject))

coded_3b = main_plus_sst %>%
  filter(assignment == "4b")
length(unique(coded_3b$subject))
```


# Analysis: Indirect requests task

After merging with the SST data, we now analyze performance on the indirect requests task.

Here, we build two logistic mixed effects models: one to predict the effect of condition ("stimType") on interpretation or knowledge, and a null model to compare the first to. 

```{r}

main_plus_sst %>%
  ggplot(aes(x = condition,
             y = numeric_categorization2)) +
  geom_bar(stat="summary", fun.y="mean", position = "dodge") +
  stat_summary (fun.y = function(x){mean(x)},
                fun.ymin = function(x){mean(x) - 1*sd(x)/sqrt(length(x))},
                fun.ymax = function(x){mean(x) + 1*sd(x)/sqrt(length(x))},
                geom= 'pointrange', 
                position=position_dodge(width=0.95)) +
  labs(x = "Condition",
       y = "Proportion of responses",
       title = "Response by condition and group") +
  theme_minimal() +
  geom_hline(yintercept = .5, linetype = "dotted") +
  facet_grid(~exp_group)


ggsave("Figures/experiment1_effect.png", 
         units="in", width=6, height=4, dpi=300)
ggsave("Figures/experiment1_effect.tiff", 
         units="in", width=6, height=4, dpi=300)
ggsave("Figures/experiment1_effect.pdf", 
       units="in", width=6, height=4, dpi=300)
```

## Descriptive

```{r}
main_plus_sst %>%
  group_by(exp_group,
           condition) %>%
  summarise(mean_request = mean(numeric_categorization2),
            sd_request = sd(numeric_categorization2))

```


## 3a

```{r}

mixed.logit = glmer(factor(categorization) ~ stimType + (1 + stimType | subject) + 
                      (1 + stimType | stimNum), 
                    data = coded_3a, 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

mixed.logit.null = glmer(factor(categorization) ~ (1  + stimType | subject) + 
                           (1 + stimType | stimNum), 
                         data = coded_3a, 
                         control=glmerControl(optimizer="bobyqa"),
                         family=binomial) 

comparison = anova(mixed.logit.null, mixed.logit)
comparison


```


### Effect of order?

```{r}
mixed.logit = glmer(factor(categorization) ~ stimType * order + 
                       (1 + stimType | subject) +
                      (1 + stimType | stimNum), 
                    data = coded_3a, 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

mixed.logit.null = glmer(factor(categorization) ~ stimType + order +
                           (1 + stimType | subject) +
                           (1 + stimType | stimNum), 
                         data = coded_3a, 
                         control=glmerControl(optimizer="bobyqa"),
                         family=binomial) 

comparison = anova(mixed.logit.null, mixed.logit)
comparison

mixed.logit = glmer(factor(categorization) ~ stimType + 
                      (1 + stimType | stimNum), 
                    data = filter(coded_3a, order ==1 ), 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

mixed.logit.null = glmer(factor(categorization) ~ (1 + stimType | stimNum), 
                         data = filter(coded_3a, order ==1 ), 
                         control=glmerControl(optimizer="bobyqa"),
                         family=binomial) 

comparison = anova(mixed.logit.null, mixed.logit)
comparison
```


## 3b

```{r}

mixed.logit = glmer(factor(categorization) ~ stimType + (1 + stimType | subject) + 
                      (1 + stimType | stimNum), 
                    data = coded_3b, 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

mixed.logit.null = glmer(factor(categorization) ~ (1  + stimType | subject) + 
                           (1 + stimType | stimNum), 
                         data = coded_3b, 
                         control=glmerControl(optimizer="bobyqa"),
                         family=binomial) 

comparison = anova(mixed.logit.null, mixed.logit)
comparison

```


### Effect of order?

```{r}
mixed.logit = glmer(factor(categorization) ~ stimType * order + 
                       (1 + stimType | subject) +
                      (1 + stimType | stimNum), 
                    data = coded_3b, 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

mixed.logit.null = glmer(factor(categorization) ~ stimType + order +
                           (1 + stimType | subject) +
                           (1 + stimType | stimNum), 
                         data = coded_3b, 
                         control=glmerControl(optimizer="bobyqa"),
                         family=binomial) 

comparison = anova(mixed.logit.null, mixed.logit)
comparison

mixed.logit = glmer(factor(categorization) ~ stimType + 
                      (1 | stimNum), 
                    data = filter(coded_3b, order ==1 ), 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

mixed.logit.null = glmer(factor(categorization) ~ (1 | stimNum), 
                         data = filter(coded_3b, order ==1 ), 
                         control=glmerControl(optimizer="bobyqa"),
                         family=binomial) 

comparison = anova(mixed.logit.null, mixed.logit)
comparison
```

## Interaction between group and main effect

```{r}
mixed.logit.interaction = glmer(factor(numeric_categorization2) ~ condition * exp_group +
                      (1 + condition | stimNum) +
                      (1 + condition | subject), 
                    data = main_plus_sst, 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

mixed.logit.both = glmer(factor(numeric_categorization2) ~ condition + exp_group +
                      (1 + condition | stimNum) +
                      (1 + condition | subject), 
                    data = main_plus_sst, 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

mixed.logit = glmer(factor(numeric_categorization2) ~ condition + 
                      (1 + condition | stimNum) +
                      (1 + condition | subject), 
                    data = main_plus_sst, 
                    control=glmerControl(optimizer="bobyqa"),
                    family=binomial) 

anova(mixed.logit.interaction, mixed.logit.both)
anova(mixed.logit.both, mixed.logit)

```




## Indirect requests and SST 


### 3a

```{r}
coded_3a$rc.bins = cut(coded_3a$rc.mean, breaks=c(0, median(coded_3a$rc.mean), 11), labels=c("low", "high"))

```


#### Analysis

```{r}

mixed.logit.both = glmer(factor(categorization) ~ stimType*exp.inf.mean +
                          stimType * rc.bins +
                          (1 + stimType  | stimNum), 
                        data = coded_3a, 
                        family=binomial) 


mixed.logit.both.reduced = glmer(factor(categorization) ~ exp.inf.mean +
                          stimType * rc.bins +
                          (1 + stimType  | stimNum), 
                        data = coded_3a, 
                        family=binomial) 

mixed.logit.both.reduced2 = glmer(factor(categorization) ~ exp.inf.mean * stimType + rc.bins +
                          (1 + stimType  | stimNum), 
                        data = coded_3a, 
                        family=binomial) 

mixed.logit.inf= glmer(factor(categorization) ~ exp.inf.mean * stimType + 
                          (1 + stimType  | stimNum), 
                          control=glmerControl(optimizer="bobyqa"),
                        data = coded_3a, 
                        family=binomial) 

mixed.logit.inf.reduced = glmer(factor(categorization) ~ exp.inf.mean + stimType +
                          (1 + stimType  | stimNum), 
                        data = coded_3a, 
                        control=glmerControl(optimizer="bobyqa"),
                        family=binomial) 

anova(mixed.logit.both, mixed.logit.both.reduced)
anova(mixed.logit.both, mixed.logit.both.reduced2)
anova(mixed.logit.inf, mixed.logit.inf.reduced)


```


```{r}
mixed.logit.spon = glmer(factor(categorization) ~ stimType*factor(spon.true) +
                          (1 + stimType | stimNum), 
                        data = coded_3a, 
                        family=binomial) 
mixed.logit.spon.reduced = glmer(factor(categorization) ~ stimType + factor(spon.true) +
                          (1 + stimType | stimNum), 
                        data = coded_3a, 
                        family=binomial) 

anova(mixed.logit.spon, mixed.logit.spon.reduced)
```


### 3b

```{r}
coded_3b$rc.bins = cut(coded_3b$rc.mean, breaks=c(0, median(coded_3b$rc.mean), 11), labels=c("low", "high"))
```



#### Analysis

```{r}
mixed.logit.both = glmer(factor(categorization) ~ stimType*exp.inf.mean +
                          stimType * rc.bins +
                          (1 + stimType  | stimNum), 
                         control=glmerControl(optimizer="bobyqa"),
                        data = coded_3b, 
                        family=binomial) 


mixed.logit.both.reduced = glmer(factor(categorization) ~ exp.inf.mean +
                          stimType * rc.bins +
                          (1 + stimType  | stimNum), 
                          control=glmerControl(optimizer="bobyqa"),
                        data = coded_3b, 
                        family=binomial) 

mixed.logit.both.reduced2 = glmer(factor(categorization) ~ exp.inf.mean * stimType + rc.bins +
                          (1 + stimType  | stimNum), 
                        data = coded_3b, 
                        control=glmerControl(optimizer="bobyqa"),
                        family=binomial) 

mixed.logit.inf= glmer(factor(categorization) ~ exp.inf.mean * stimType + 
                          (1 + stimType  | stimNum), 
                          control=glmerControl(optimizer="bobyqa"),
                        data = coded_3b, 
                        family=binomial) 

mixed.logit.inf.reduced = glmer(factor(categorization) ~ exp.inf.mean + stimType +
                          (1 + stimType  | stimNum), 
                        data = coded_3b, 
                        control=glmerControl(optimizer="bobyqa"),
                        family=binomial) 

anova(mixed.logit.both, mixed.logit.both.reduced)
anova(mixed.logit.both, mixed.logit.both.reduced2)
anova(mixed.logit.inf, mixed.logit.inf.reduced)
```




```{r}
mixed.logit.spon = glmer(factor(categorization) ~ stimType*factor(spon.true) +
                          (1 + stimType | stimNum), 
                        data = coded_3b, 
                        family=binomial) 
mixed.logit.spon.reduced = glmer(factor(categorization) ~ stimType + factor(spon.true) +
                          (1 + stimType | stimNum), 
                        data = coded_3b, 
                        family=binomial) 

anova(mixed.logit.spon, mixed.logit.spon.reduced)
```


### Visualizations

```{r}
main_plus_sst %>%
  ggplot(aes(x=exp.inf.mean, fill=Congruence)) +
  geom_density(alpha=.5) +
  theme_minimal() +
  scale_fill_manual(values = c("congruent" = "seagreen2", "incongruent" = "tomato1")) +
  xlab("Mentalizing score") +
  ggtitle("Response congruence by mentalizing") +
  facet_wrap(~exp_group, ncol = 2)

ggsave("figures/exp4_mentalizing.png", 
       units="in", width=6, height=4, dpi=300)
ggsave("figures/exp4_mentalizing.pdf", 
       units="in", width=6, height=4, dpi=300)
ggsave("figures/exp4_mentalizing.tiff", 
       units="in", width=6, height=4, dpi=300)
```

```{r}
main_plus_sst %>%
  ggplot(aes(x=exp.inf.mean, fill=categorization)) +
  geom_density(alpha=.5) +
  theme_minimal() +
  # scale_fill_manual(values = c("congruent" = "seagreen2", "incongruent" = "tomato1")) +
  xlab("Mentalizing score") +
  ggtitle("Response congruence by mentalizing") +
  facet_grid(~assignment + condition)
```


### Mentalizing three-way interaction


```{r}
main_plus_sst$rc.bins = cut(main_plus_sst$rc.mean, 
                            breaks=c(0, median(main_plus_sst$rc.mean), 11), 
                            labels=c("low", "high"))
mixed.logit.both = glmer(factor(numeric_categorization2) ~ condition*exp.inf.mean*exp_group +
                          stimType * rc.bins +
                          (1 + stimType  | stimNum), 
                         control=glmerControl(optimizer="bobyqa"),
                        data = main_plus_sst, 
                        family=binomial) 

mixed.logit.reduced = glmer(factor(numeric_categorization2) ~ condition*exp.inf.mean +
                           condition*exp_group +
                           exp.inf.mean*exp_group+
                          stimType * rc.bins +
                          (1 + stimType  | stimNum), 
                         control=glmerControl(optimizer="bobyqa"),
                        data = main_plus_sst, 
                        family=binomial) 


anova(mixed.logit.both, mixed.logit.reduced)


```

